\section{Introduction}
\epigraph{\emph{It would be hung in the attics, she thought; it would be destroyed. But what did that matter? she asked herself, taking up her brush again.}}{---Virginia Woolf, \emph{To the Lighthouse}}

\noindent Associated to any based topological space $(X,x)$ is its fundamental group $\pi_1(X,x)$. In the seminal 1895 paper \emph{Analysis Situs} \cite{poincare1895analysis}, Henri Poincar{\'e} introduced the fundamental group alongside other important concepts such as homology and cobordism, essentially founding the modern study of algebraic topology. This thesis is broadly concerned with the relationship between the fundamental group and homology: the former contains information that the latter cannot see. 
One can ask whether it is possible to learn about the fundamental group given information about homology. 

While the integral homology groups are always abelian, the fundamental group is often not. So let us instead examine the integral group ring $\Z\pi_1(X,x)$. This is unfortunately an unwieldy object: if $\pi_1(X,x)$ is infinite, which is often the case, then $\Z\pi_1(X,x)$ is of infinite rank. To turn it into something feasible to study, let us consider the quotients of $\Z\pi_1(X,x)$ by the $n$th powers of its augmentation ideal $\mathcal{I}$. A result of Stallings (Theorem \ref{thm:stallings} here, Theorem 5.4 in \cite{Stallings1975}) relates this truncated fundamental group ring to the $E^\infty$-page of a certain spectral sequence associated to the \emph{cobar construction} on the singular chains of $X$. The cobar construction is a functor sending differential graded (dg) conilpotent coalgebras to dg algebras. While it can be defined purely algebraically, it has a nice topological interpretation when performed on the singular chains of a topological space. It is through this that we will relate $\Z\pi_1(X,x)/\mathcal{I}^n$ to the homology of $X$. 
\subsection{Main theorem}
Let $X$ be a path-connected topological space and choose basepoints $a, b \in X$. Let $\pi_X(a,b)$ be the set of homotopy classes of paths in $X$ from $a$ to $b$. In particular when $a=b$ we obtain the fundamental group of $X$. Notice that $\pi_1(X,a)$ acts on $\pi_X(a,b)$ by path concatenation, making the free abelian group $\Z\pi_X(a,b)$ into a right $\Z\pi_1(X,a)$-module. Consider the following subspace:
\[X(n)^a_b = \{ (x_1,\dots,x_n) \in X^n \mid x_1 = a \;\text{or}\; x_n = b \;\text{or}\; x_i = x_{i+1} \text{ for some } i \} \subset X^n.\]
Our main theorem is the following.
\begin{thm}\label{thm:looijenga}
There is a surjective group homomorphism 
\[\Z\pi_X(a,b)/ \Z\pi_X(a,b)\mathcal{I}^{n+1}_a\to H_n(X^n, X(n)^a_b; \Z)\]
whose kernel is isomorphic to $\Z$ if $a=b$ and zero otherwise.
\end{thm}
A dual cohomological version of this result with field coefficients is proved by Deligne and Goncharov in \cite{DELIGNE2005}, Proposition 3.4, where they attribute it to Beilinson. A version for $X$ a real manifold and de Rham cohomology is developed by Chen in \cite{Chen1973}. The version stated here comes from a paper of Looijenga \cite{Looijenga2025}, Theorem 1.1, where the proof is quite concise and elegant, but (perhaps to a student) somewhat opaque. Returning to the paper of Deligne and Goncharov, in the discussion following the statement and proof of this result, the authors suggest another possible proof in terms of cosimplicial spaces. We find this avenue more enlightening. In particular, it offers an explanation for why the subspace $X(n)^a_b$ appears. 

The goal of this thesis will be to develop the necessary background to pursue this alternate proof, while hopefully clearly explicating the details of each step. Of course, the authors listed above are interested in this theorem for deeper reasons beyond just its statement and proof. We do not claim to understand why, but give a few remarks on this matter at the very end.

\subsection{Organization} In \S\ref{chapter:coalgebras} we give a crash course on the theory of coalgebras. In \S\ref{chapter:cosimplicial} we develop the basic theory of (co)simplicial objects, in particular cosimplicial topological spaces. We also collect some helpful technical lemmas about (co)simplicial objects. The reader should feel free to skip these first two sections if they are familiar with the theory. Both coalgebras and (co)simplicial objects are crucial in the following \S\ref{chapter:adamscobar}, where we introduce the cobar construction and Adams' theorem, and define a generalized two-sided version of the cobar construction. In \S\ref{chapter:stallings} we generalize Stallings' theorem to this two-sided version, and finally use all these ingredients to prove our main theorem in \S\ref{chapter:finalproof}.

I have tried to make the thesis readable to anyone who has had slightly more than a course or two in algebraic topology, and working knowledge of (homological) algebra. Here are some prerequisites:
\begin{itemize}
\item Algebraic topology at the level of the first two chapters of Hatcher's book \cite{Hatcher:478079}.
\item Homological algebra, in particular the spectral sequences associated to a bicomplex. They unfortunately play a key computational role throughout the thesis. The proof of the main theorem comes from the close examination of one particular spectral sequence. We have included proofs of some auxillary results about these spectral sequences, but the reader could possibly regard them as a black box.
\item Basic notions in category theory: functors, natural transformations, adjunctions, (co)limits. 
\item Some prior exposure to simplicial sets, e.g. the singular-geometric realization adjunction. This will be very helpful for intuition with cosimplicial objects even though we technically develop the theory of cosimplicial objects from scratch. Simplicial sets will also feature heavily in \S\ref{chapter:stallings}, though we really only care when the simplicial set in question is $\Sing(X)$ for a topological space $X$.
\end{itemize}

A few last words on convention. (1) Often we are working with a bicomplex $E$ with a totalization $\tot E$. We will often denote the homology of the totalization by just $H_\bullet E$ rather than $H_\bullet\tot E$. (2) The letter $R$ is reserved for commutative unital rings. Other rings appearing may not be commutative nor unital. (3) Normalized or singular chains of a topological space will have coefficients in $\Z$. (4) If $Y$ is a simplicial set then $|Y|$ denotes its geometric realization. (5) ``Space'' means topological space. (6) All spectral sequences will be second quadrant, and depending on the filtration, the $r$th differential will have bidegree $(r, 1-r)$ or $(1-r, r)$. 